\pkg{DNest4}: Diffusive Nested Sampling in \proglang{C++} and \proglang{Python}

### 2016-05-23 paper/ms.tex
More writing

\begin{CodeChunk}
\begin{CodeInput}
# File containing parameters for DNest4
# Put comments at the top, or at the end of the line.
5       # Number of particles
10000   # new level interval
10000   # save interval
100     # threadSteps - pooling interval
0       # maximum number of levels (0 = auto)
10      # Backtracking scale length (lambda in the paper)
100     # Strength of effect to force histogram to equal push. Beta in the paper
5000    # Maximum number of saves (0 = infinite)
\end{CodeInput}
\end{CodeChunk}

The first option is the number of particles, here set to five.
If you use more particles, the same amount of CPU time will be spent evolving more particles,
so each one will not be evolved as far. On most problems, five is a sensible
default value. On complex problems where the likelihood function has
a challenging structure, more particles are useful, but it is usually better
%to run in multi-threaded mode (see Appendix~\ref{sec:command_line_options}).

The new level interval controls how quickly \dnest~creates new levels. In this
example, this is set to 10,000, so a new level will be created once \dnest~has
seen 10,000 likelihood values above the current top level. This is not affected by
the number of threads. It is difficult to give a sensible default for this
quantity because it depends on the complexity of the problem and how good
the Metropolis proposals are. However, 10,000
will work for many problems (I rarely use anything less than 10,000 or greater
than 300,000). Higher values are slower, but more fail-safe.

The save interval controls how often \dnest~writes a model to the output
files: this is what is usually called ``thinning''. Saving more frequently
(i.e. a smaller save interval) is usually better, although it depends on your
problem. If your models have a lot of parameters {\tt sample.txt} can get
very big and take forever to process using {\tt showresults.py}. A default
suggestion that is usually okay is to make the save interval the same as the
new level interval.

>>>>

\begin{CodeChunk}
\begin{CodeInput}
# File containing parameters for DNest4
# Put comments at the top, or at the end of the line.
5       # Number of particles
10000   # new level interval
10000   # save interval
100     # threadSteps - pooling interval
0       # maximum number of levels (0 ==> automatic)
10      # Backtracking scale length (lambda in the paper)
100     # Equal weight enforcement. Beta in the paper
5000    # Maximum number of saves (0 ==> run forever)
\end{CodeInput}
\end{CodeChunk}

The first option is the number of particles, here set to five.
If you use more particles, the same amount of CPU time will be spent evolving more particles,
so each one will not be evolved as far. On most problems, five is a sensible
default value. On complex problems where the likelihood function has
a challenging structure, more particles are useful, but it is usually better
to run in multi-threaded mode (see Appendix~\ref{sec:command_line_options}).

The new level interval controls how quickly \dnest~creates new levels. In this
example, this is set to 10,000, so a new level will be created once 10,000
MCMC steps have resulted in
10,000 likelihood values above the current top level.
It is difficult to give a sensible default for this
quantity because it depends on the complexity of the problem (basically,
how good the Metropolis proposals are at exploring the target distribution).
However, 10,000 will work for many problems, so we suggest it as a sensible
default. Higher values are slower, but more fail-safe.

The save interval controls how often \dnest~writes a model to the output
files; what is usually called ``thinning''. Saving more frequently
(i.e. a smaller save interval) is usually better. However, this can result
in
big output files if your model prints a lot of parameters to {\tt sample.txt}.
causing the postprocessing to take a long time and/or a lot of RAM.
A default suggestion, used in the example, is to set the save interval to the
same value as the new level interval.
